### 2023-07-11 paper/paper.tex
[ paper ] various writing

\documentclass[acmsmall,review,screen]{acmart}
\settopmatter{printfolios=true,printccs=true,printacmref=true}

\usepackage{cmll}
\usepackage{mathpartir}

\newcommand{\TapeType}{\mathrm{Tape}}

\newcommand{\tmRec}{\mathrm{rec}}
\newcommand{\tyNat}{\mathrm{Nat}}
\newcommand{\conZero}{\mathsf{zero}}
\newcommand{\conSucc}{\mathsf{succ}}
\newcommand{\conRefl}{\mathsf{refl}}

\newcommand{\Let}{\mathrm{let}}
\newcommand{\LetPair}{\mathrm{letpair}}
\newcommand{\In}{\mathrm{in}}
\newcommand{\If}{\mathrm{if}}
\newcommand{\Then}{\mathrm{then}}
\newcommand{\Else}{\mathrm{else}}
\newcommand{\dupNat}{\mathrm{dupNat}}
\newcommand{\Match}{\mathrm{match}}

% QTT stuff
\newcommand{\istype}{\mathrm{type}}
\newcommand{\isctxt}{\mathrm{ctxt}}

\newcommand{\BoolTy}{\mathrm{Bool}}
\newcommand{\cTrue}{\mathsf{true}}
\newcommand{\cFalse}{\mathsf{false}}
\newcommand{\ListTy}{\mathrm{List}}
\newcommand{\cNil}{\mathsf{nil}}
\newcommand{\cCons}{\mathsf{cons}}

\newcommand{\Rtype}{\mathbf{R}}
\newcommand{\rIntro}{\mathbf{R}}
\newcommand{\rElim}{\mathbf{R}^{-1}}

\newcommand{\natinf}{\mathbb{N}_{-\infty}}

\newcommand{\Ty}{\mathrm{Ty}}
\newcommand{\RTm}{\mathrm{RTm}}
\newcommand{\Tm}{\mathrm{Tm}}

\newcommand{\Set}{\mathrm{Set}}

\newcommand{\cat}[1]{\mathcal{#1}}
\newcommand{\op}{\mathsf{op}}

\newcommand{\LinPreorder}{\mathrm{LinPreorder}}

\newcommand{\bob}[1]{\textcolor{purple}{FIXME: #1}}

\renewcommand{\sectionautorefname}{Section}
\renewcommand{\subsectionautorefname}{Section} % PS
\renewcommand{\subsubsectionautorefname}{Section}

\newcommand{\lemref}[1]{\hyperref[#1]{Lemma \ref*{#1}}}
\newcommand{\exref}[1]{\hyperref[#1]{Example \ref*{#1}}}
\newcommand{\judgeref}[1]{\hyperref[#1]{Judgement \ref*{#1}}}
\newcommand{\propref}[1]{\hyperref[#1]{Proposition \ref*{#1}}}
\newcommand{\diagref}[1]{\hyperref[#1]{Diagram \ref*{#1}}}
\newcommand{\thmref}[1]{\hyperref[#1]{Theorem \ref*{#1}}}
\newcommand{\defref}[1]{\hyperref[#1]{Definition \ref*{#1}}}

\newcommand{\AgdaModule}[1]{\textcolor{blue}{\tt #1}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}
\acmDOI{10.1145/1122445.1122456}

%%
%% These commands are for a JOURNAL article.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{CONF}
\acmArticle{1}
\acmMonth{8}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Polynomial Time and Full Dependent Types}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Robert Atkey}
\email{robert.atkey@strath.ac.uk}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{University of Strathclyde}
  \streetaddress{26 Richmond Street}
  \city{Glasgow}
  \country{UK}
  \postcode{G1 1XH}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  We combine dependent types with linear type systems that soundly and
  completely capture polynomial time computation. To combine dependent
  and linear types, we use \emph{Quantitative Type Theory}. We explore
  two systems for capturing polynomial time: one system that disallows
  construction of iterable data, and one, based on the LFPL system of
  Martin Hofmann, that controls construction via a payment
  method. Both of these are extend to full dependent types, allowing
  for arbitrary computation in types alongside guaranteed polynomial
  time computation in terms.

  The resulting system smoothly combines reasoning with dependent
  types and restricted computational complexity. \bob{Elaborate after
    Section 3 is written.}We prove the soundness of the systems using
  a realisability technique due to Hofmann and Dal Lago.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{type theory, implicit computational complexity, linear logic, computational complexity}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}

{\bf New story:}
\begin{enumerate}
\item Introduction: The goal of this paper is to unify implicit
  computational complexity with full dependent types. ICC enables us
  to ensure that \emph{programs} are within a certain complexity
  class, while dependent types allow us to state and prove that these
  programs solve the problems that we intend them to, and to construct
  reductions between problems. We can marry complexity-controlled
  implementations with problems, paving the way for a \emph{synthetic
    computational complexity theory}.
\item Section 1: Linear types and Implicit Polynomial Time.
  \begin{enumerate}
  \item Total typed functional programming, as we see in type theory,
    is restricted to the use of higher order functions and iteration
    over inductive data types.
  \item Polynomial time is the smallest class closed under nested
    iteration over the input -- i.e. we consider iteration over the
    input as feasible, and compositions thereof. The key thing is to
    avoid iteration over data constructed from the input.
  \item Key is to control iteration. So we split inductive datatypes
    into iterable and non-iterable versions. We can
  \item One way to capture polytime is to ensure that all iteration is
    only done over the input. So we can duplicate iterable things, but
    we can't construct new iterable things (at least not from an
    iteration). We also can't duplicate higher order things, because
    that effectively acts as a way to duplicate iterations (half
    instantiated iterators). Compare to Jones' CONS-FREE programming.
  \item Another approach is to allow ``borrowing'' of iterability by
    means of diamonds. Iterability becomes a first-class object, so we
    can use it to construct new data structures from the input. This
    allows the construction of new iterable data structures. LFPL.
  \end{enumerate}
\item Section 2: Extending to Dependent Types
  \begin{enumerate}
  \item Combining linear types and dependent types is subtle, because
    there is a fundamental incompatibility in the syntax of the two
    systems.
  \item There have been several approaches, though they are all
    somewhat similar (list them all here). We will use QTT, because it
    offers a firm semantic foundation as well as a relatively
    straightforward syntax.
  \item Recap the syntax of QTT, with extensional equality. Key point
    is that there is a normal dependently typed language with a linear
    sub-language. We can have special constructs in the linear
    sub-language that ``erase'' to normal type theory -- products are
    the prime example.
  \item Non iterable datatypes -- have normal eliminators in the DTT
    fragment, but
  \end{enumerate}
\item Section 3: Programming and Proving with Poly-time
  \begin{enumerate}
  \item Internalised completeness proofs.
  \item Construct the type of problems that are solvable in polynomial
    time.
  \item Construct reductions between problems.
  \item NP. Doesn't make sense to talk about this without dependent
    types, because we can't say when we've solved the problem being
    stated.
  \item PP (Probabilistic), using a monad. Again, need to be able to
    state connection with the problem to know that we are solving it.
  \end{enumerate}
\item Section 4: Soundness via Realisability
  \begin{enumerate}
  \item The underlying CBV language and its
  \end{enumerate}
\item Section 5: Extensions
  \begin{enumerate}
  \item Explicit Complexity via Diamond Counting
  \item Poly-time univalence
  \item Towards P- and NP-completeness proofs. Issue here is the
    extensionality. Towards a synthetic (poly-time based)
    computational complexity theory.
  \end{enumerate}
\item Section 6: Related Work
  \begin{enumerate}
  \item ICC stuff
    \begin{enumerate}
    \item LFPL
    \item CONS-free
    \item Soft Affine Logic; which has the same underlying idea as
      CONS-free (is this a novel observation?)
    \item Other systems based on information flow.
    \end{enumerate}
  \item Explicit complexity: lambda-amor, and things by Deepak Garg
    and Marco Gaboardi.
  \item Recent CMU stuff on calf
  \end{enumerate}

\end{enumerate}

{\bf Old story:}
\begin{enumerate}
\item There are systems for implicit computational complexity based on
  linear logic (e.g. LFPL, SAL, LLL, BLL) that seek to capture
  complexity classes (most often polynomial time) implicitly
  (i.e. without directly talking about time bounds). This work has
  also led to followup work on explicitly resourced calculi such as
  RAML, which seek to use static analysis to determine complexity
  bounds on programs.
\item In dependent type theory, we can reason about the extensional
  behaviour of programs (within the limits of intensional type
  theory), but not about intenstional properties. The only way to
  formalise something like ``for all polynomial time functions
  $\mathbb{N} \to \mathbb{N}$'' is to develop a deep emebdding of
  polynomial time functions and to work with that, which is tedious.
\item The goal of this paper is to unify dependent types for reasoning
  about programs with implicit computational complexity type
  systems. In the first instance, we do not seek to use dependent
  types to state precise bounds on programs' time consumption, but we
  intend to use dependent types to reason about the extensional
  behaviour of programs that we know to be polynomial time. At the end
  of the paper, we show how to adapt our framework to explicit
  resource accounting, in the style of RAML.
\item The plan of the paper is as follows:
  \begin{enumerate}
  \item We first introduction the ideas behind the use of linear logic
    for capturing polynomial time, by studying two systems Lafont's
    SAL, and Hofmann's LFPL. In Hofmann and Dal Lago's presentation,
    these are two variants of Second order Multiplicative Linear
    Logic, with restricted $\oc$ modalities for duplication. We also
    recall the realisability semantics of Dal Lago and Hofmann (FIXME:
    move this later?)
  \item The most modern presentations of these systems use second
    order quantification to represent data types. This is incompatible
    with the form of dependent types we use here, so we seek a direct
    representation of natural numbers that captures
  \item We then recall QTT, a extension of Martin-Löf type theory with
    linear types, which can accomodate intensional restrictions on
    computability as well as full reasoning about programs. Since
    allowing an unrestricted $\oc$ modality would violate our
    polynomial time bounds, we instead choose the natural number
    semiring for resource accounting.
  \item We then transplant the simply typed rules into QTT, showing
    how to carefully distinguish between data that is available to the
    program, and data that is only available for type formation.
  \item We finish by showing that we can also do explicit resource
    tracking.
  \end{enumerate}
\end{enumerate}

Related work: only applications of Linear TT so far are erasure (Abel
et al.) and quantum stuff (Selinger et al.), and (sort of) dependency
analysis. This is another one.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Linear $\lambda$-Calculus and Polytime}
\label{sec:linear-intro}

Not long after Girard introduced Linear Logic \cite{girard87}, it was
observed that its resource sensitivity could be turned to describing
computational complexity classes by purely logical means. Typically, a
logical system is described for which the process of reducing a proof
to a normal form (often by cut elimination) is guaranteed to always be
accomplished within a certain complexity bound. Moreover, the system
is usually proven to be complete for the relevant complexity class, by
constructing a simulation of some known representation. Such systems
that characterise polytime include Bounded Linear Logic (BLL), which
uses explicit polynomials in the formulas \cite{bll} and Soft Affine
Logic (SAL) \cite{lafont}, which does not explicitly represent
time information in formulas, but uses a restricted form of Linear
Logic's $\oc$ modality instead. Light Affine Logic (LAL) \cite{lal} is
another ``counting-free'' system for polytime.

Viewing logical systems though the lens of the Curry-Howard
correspondence, the idea arises that one could define functional
programming languages that characterise complexity classes such as
polytime. SAL has been transformed into a programming language for
polytime by \cite{gaboardi}. Hofmann proposed a new programming
language, Linear Functional Programming Language (LFPL), that uses a
novel ``payment'' system to track iteration. In this section, we
review the use of linear types to capture polytime by presenting two
systems, one based on ideas from SALL and the second more explicitly
based on LFPL. With a view to adapting these systems to dependent
types in \autoref{sec:qtt}, we take an approach slightly different to
much of the polytime linear logic literature. We emphasise the use of
explicit datatypes and eliminators, rather than using impredicative
encodings via universal types. Thus, we are closer to Hofmann's
original LFPL (though not later presentations of it
\cite{dallago-hofmann}) than BLL, SALL, or LAL.

For this section, the linear $\lambda$-calculus we will use will have
linear functions and $\otimes$-products:
\begin{mathpar}
  \inferrule*
  { }
  {x : A \vdash x : A}

  \inferrule*
  {\Gamma, x : A \vdash M : B}
  {\Gamma \vdash \lambda x.M : A \multimap B}

  \inferrule*
  {\Gamma_1 \vdash M : A \multimap B \\
    \Gamma_2 \vdash N : A}
  {\Gamma_1, \Gamma_2 \vdash M\,N : B}

  \inferrule*
  {\Gamma_1 \vdash M : A \\ \Gamma_2 \vdash N : B}
  {\Gamma_1, \Gamma_2 \vdash (M,N) : A \otimes B}

  \inferrule*
  {\Gamma_1 \vdash M : A \otimes B \\ \Gamma_2, x : A, y : B \vdash N : C}
  {\Gamma_1, \Gamma_2 \vdash \Let\,(x,y)=M\,\In\,N : C}
\end{mathpar}
These rules are standard, so we do not describe them further except to
note how linear typing uses presence or absences in a context to
control resource usage -- if a variable is in the context it must be
used exactly once. The fact that this discipline interferes with
dependent types is one of the reasons we turn to QTT when we wish to
add dependent types in \autoref{sec:qtt}.

\subsection{No Recursion, Only Case Analysis}
\label{sec:linear-case-analysis}

It is not too difficult to see that reduction of linear
$\lambda$-terms always takes a number of steps linearly proprotional
to the size of the term. This is because every $\beta$-redex
substitutes into exactly one variable, reducing the size of the term
by one each time.

We can increase the expressivity, but not the computational
complexity, of the system by adding datatypes that do not allow
iteration. These can be used for representation, but not for driving
computation. We include the rules here, to show how linearity must be
preserved in these rules, and to foreshadow their dependently typed
counterparts in \autoref{sec:qtt-noniter}. The first type is the
booleans, which are anyway non-recursive and so wouldn't allow
iteration anyway:
\begin{mathpar}
  \inferrule*
  { }
  {\vdash \cTrue, \cFalse : \BoolTy}

  \inferrule*
  {\Gamma_1 \vdash M : \BoolTy \\ \Gamma_2 \vdash N_1 : A \\ \Gamma_2 \vdash N_2 : A}
  {\Gamma_1, \Gamma_2 \vdash \If\,M\,\Then\,N_1\,\Else\,N_2 : A}
\end{mathpar}
The if-then-else rule is careful to ensure that the resources used by
the eliminated $\BoolTy$ and the resources used by the chosen branch
are accounted for separately. The two branches must have the same
resource usage.

Construction and case analysis of lists are given by the following
rules:
\begin{mathpar}
  \inferrule*
  { }
  {\vdash \cNil : \ListTy(A)}

  \inferrule*
  {\Gamma_1 \vdash M : A \\ \Gamma_2 \vdash N : \ListTy(A)}
  {\Gamma_1, \Gamma_2 \vdash \cCons(M,N) : \ListTy(A)}

  \inferrule*
  {\Gamma_1 \vdash M : \ListTy(A) \\
    \Gamma_2 \vdash N_1 : B \\
    \Gamma_2, h : A, t : \ListTy(A) \vdash N_2 : B}
  {\Gamma_1, \Gamma_2 \vdash \Match\,M\,\{\cNil \mapsto N_1; \cCons(h,t) \mapsto N_2\} : B}
\end{mathpar}
Hence, we can construct lists arbitrarily, but we can only do case
analysis on them. If we wish to explore a list to a arbitrary depth it
must be driven by an iterable type of the form that we will see in the
next two sections.

With booleans and lists, we can construct several other useful
types. For example, to simulate Turing machines, one can construct a
$\mathit{Tape}$ type as a Zipper \cite{huet}
$\ListTy(\BoolTy) \otimes \BoolTy \otimes \ListTy(\BoolTy)$,
representing a position on the tape with the items before, under, and
after the head.

\subsection{The Cons-free system}
\label{sec:cons-free-intro}

Polynomial time is usually seen as a proxy for ``feasible''
computation. On the face of it, there does not seem to be any
particular reason why polynomials have anything to do with
feasibility. However, one can arrive at the definition of polynomial
time in three steps, by assuming that (i) iterating over the whole
input is feasible; (ii) if two computations are feasible, then so is
their composition; and (iii) performing a feasible computation for
every element of the input is also feasible. It is the last point that
allows complexities of arbitrary polynomial degree to be constructed
(we will see this in action in the completeness construction below and
soundness proofs in \autoref{sec:iterator-soundness}).

Following these ideas, let us assume that the input is a natural
number, so we assume that there is some type of natural numbers
$\tyNat$. For point (i), we must be able to iterate over these natural
numbers, so we use a linear iterator defined by this typing rule:
\begin{displaymath}
  \inferrule*
  {\vdash M_z : A \\ x : A \vdash M_s : A \\ \Gamma \vdash N : \tyNat}
  {\Gamma \vdash \tmRec\,N\,\{\conZero \mapsto M_z; \conSucc(x) \mapsto M_s\} : A}
\end{displaymath}
Note that in the zero, $M_z$, and sucessor, $M_s$, cases, the context
is empty to ensure that these cases may be invoked as many times as
required. Point (ii) above is automatically satisfied by being in a
typed $\lambda$-calculus, where it is difficult to stop functions from
being composable. For point (iii), the iterator as given does not
allow us to nest iterations. Once the natural number input $n$ has
been used for an iteration, the linear typing discipine prevents us
from using it again (note the two separate contexts $\Gamma_1$,
$\Gamma_2$ in the rule for application). In order to allow nested
iterations, we add an operator to duplicate numbers:
\begin{displaymath}
  \inferrule*
  {\Gamma \vdash M : \tyNat}
  {\Gamma \vdash \dupNat\,M : \tyNat \otimes \tyNat}
\end{displaymath}
Somewhat surprisingly, this system is now sound and complete for
polynomial time. Crucially, this depends on the two things we have
\emph{not} allowed. First, we have disallowed the construction of new
natural numbers via the $\conZero$ and $\conSucc$
constructors\footnote{Actually, $\conZero$ would be acceptable, as
  well as any constant natural number. It is only unrestricted use of
  $\conSucc$ that is dangerous.}. If we were to permit this, then we
could use iteration over the input to construct addition,
multiplication (by repeated addition), and then exponentials (by
repeated multiplication). We therefore refer to this system as the
\emph{Cons-Free} system.  The second prohibited feature is the ability
to duplicate values of function type, even though we have allowed
duplication of iterable naturals. If we were to allow this, then we
would be able to sneak in a form of constructors for natural numbers
by encoding them as eliminators that duplicate a function for every
$\conSucc$ step.

We will see in \autoref{sec:consfree-sound} that this system is sound
for polytime by a realisability argument. Completeness can be seen
more directly by constructing a function that iterates a function for
a statically known polynomial number of times in the size of the
input. Assume that we have a known polynomial
$p(n) = c_dn^d + \dots + c_0$ of degree $d$ with natural number
coefficients and some single step function
$f : \mathit{St} \multimap \mathit{St}$ over a state type
$\mathit{St}$ that runs to completion for input of size $n$ in $p(n)$
steps. Then, using the iterator above we can iterate $f$ over a
$\tyNat$ representing the size of the input:
\begin{displaymath}
  \begin{array}{l}
    I_1 : \tyNat \multimap \mathit{St} \multimap \mathit{St} \\
    I_1 = \lambda n. \lambda s.\tmRec\,n\,\{\conZero \mapsto s; \conSucc(s) \mapsto f\,s \}
  \end{array}
\end{displaymath}
To achieve higher degrees, we can use $\dupNat$ to nest iterations:
\begin{displaymath}
  \begin{array}{l}
    I_{k+1} : \tyNat \multimap \mathit{St} \multimap \mathit{St} \\
    I_{k+1} = \lambda n. \lambda s.
    \begin{array}[t]{@{}l}
      \Let\,(n,n') = \dupNat\,n\,\In\\
      \tmRec\,n\,\{\conZero \mapsto s; \conSucc(s) \mapsto I_k\,n'\,s \}
    \end{array}
  \end{array}
\end{displaymath}
By further use of $\dupNat$ and composition to handle addition of
polynomials, the function $f$ can now be iterated $p(n)$ many times,
where $n$ is the input $\tyNat$. Thus, the \emph{Cons-free} system can
represent all polytime computations.

\subsection{Diamond Trading with LFPL}
\label{sec:lfpl-intro}

The Cons-free system is sound and complete for polytime, but is quite
awkward from the point of view of functional programming. It allows us
to iterate over natural numbers that come from the input but does not
allow us to build further values to do iteration on. For example, if
our input is a list, then we cannot transform it into a binary search
tree and then flatten it, we must always refer back to the original
natural number input. Even dividing the input into two halves to be
treated separately is difficult.

A more flexible system was proposed by \cite{hofmann99}. The
\emph{Linear Functional Programming Language} (LFPL) takes a different
approach to the \emph{Cons-free} system. Instead of completely
prohibiting construction of new data, constructions is allowed if it
is paid for by values of new type $\Diamond$ (``diamonds''):
\begin{mathpar}
  \inferrule*
  {\Gamma \vdash M : \Diamond}
  {\Gamma \vdash \conZero(M) : \tyNat}

  \inferrule*
  {\Gamma_1 \vdash M : \Diamond \\ \Gamma_2 \vdash N : \tyNat}
  {\Gamma_1, \Gamma_2 \vdash \conSucc(M, N) : \tyNat}
\end{mathpar}
Thus, to construct a $\conZero$, we must have a $\Diamond$ to pay for
it, and likewise, to construct a $\conSucc$ we must have a diamond. We
can think of $\Diamond$s as an unit of iterable data. Iterability is
``saved up'' in data during construction, and released during
iteration. Diamonds cannot be created by a program itself, for the
same reason that constructors were prohibited in the \emph{Cons-free}
system, but they are released from iterable data during iteration. The
LFPL natural number iterator has the following typing rule:
\begin{displaymath}
  \inferrule*
  {d : \Diamond \vdash M_z : A \\ d : \Diamond, x : A \vdash M_s : A \\ \Gamma \vdash N : \tyNat}
  {\Gamma \vdash \tmRec\,N\,\{\conZero(d) \mapsto M_z; \conSucc(d,x) \mapsto M_s \} : A}
\end{displaymath}
The difference with the \emph{Cons-free} iterator above is that the
$\conZero$ and $\conSucc$ cases now both have an addition binding of
type $\Diamond$. This allows some form nesting of iterations: during
an iteration over the input, the program can accumulate $\Diamond$s to
use for iteration over substructures that are smaller than the current
point in the iteration. A construction, due to Aelhig and
Schwictenberg \cite{syntactic-lfpl}, illustrates how this leads to all
polytime computations. As above, we assume a polynomial $p(n)$ and a
step function $f : \mathit{St} \multimap \mathit{St}$ that needs to be
iterated $p(n)$ times. We construct a linear iterator:
\begin{displaymath}
  \begin{array}{l}
    I_1 : (\tyNat \otimes \mathit{St}) \multimap (\tyNat \otimes \mathit{St}) \\
    I_1 = \lambda (n, s).\,\tmRec\,n\,\{
    \begin{array}[t]{@{}lcl}
      \conZero(d)&\mapsto&(\conZero(d), s);\\
      \conSucc(d,(n,s)) &\mapsto& (\conSucc(d,n),f\,s) \}
    \end{array}
  \end{array}
\end{displaymath}
Note that this iterator returns the natural number input as well as
the new state. LFPL does not allow duplication of iterable inputs, so
we must always reconstruct it if we want to do further
iteration. Addition of polynomials is accomplished by composition of
iterators. To raise the degree, we again use a nesting iterator:
\begin{displaymath}
  \begin{array}{l}
    I_{k+1} : (\tyNat \otimes \mathit{St}) \multimap (\tyNat \otimes \mathit{St}) \\
    I_{k+1} = \lambda (n,s).\,\tmRec\,n\,\{
    \begin{array}[t]{@{}lcl}
      \conZero(d)&\mapsto&(\conZero(d),s); \\
      \conSucc(d,(n,s)) &\mapsto& \Let\,(n,s)=I_k\,(n,s)\,\In\,(\conSucc(d,n), f\,s) \}
    \end{array}
  \end{array}
\end{displaymath}
Unlike in the \emph{Cons-free} system, this iterator does not raise
the degree of the nested iterator directly. Rather the iterator $I_k$
performs $\binom{n}{k}$ iterations. As observed by Aelhig and
Schwictenberg, this is sufficient because the binomials form a basis
for the vector space of all polynomials.

Despite this slightly more involved completeness construction, the
advantage of LFPL is that it is now easy to have arbitrary iterable
datatypes and to transform between them. We need only take the
introduction and elimination rules for any inductive datatype and add
$\Diamond$ premises to the introduction rules and $\Diamond$ bindings
to the eliminators.

% \subsection{Explicit Resource Tracking}

% \begin{enumerate}
% \item We still require diamonds to build zero and successor, but the
%   iterator does not give them back.
% \item In the simply typed case, we can annotate types with the static
%   amount of potential they have per constructor.
% \end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Polytime Quantitative Type Theory}
\label{sec:qtt}

We have now seen the \emph{Cons-free} and \emph{LFPL} systems for
capturing polytime by means of linear typing and restricted
iteration. We now look to extend these systems to include dependent
types.

To integrate linear typing for controlling time complexity with
dependent types we use \emph{Quantitative Type Theory} (QTT)
\cite{atkey18qtt}. We now review the definition of QTT and describe
how we have adapted it to the polytime systems we saw in the previous
section.

\subsection{Quantitative Type Theory}
\label{sec:qtt-basic}

Integrating linear and dependent types is not straightforward due to
the conflict between the linear typing discipline regarding presence
of a variable as only bestowing the right to use it once, and the
dependent typing regime that uses variables both in types (for
specification purposes) and in terms (for computational purposes),
syntactically yield multiple uses of the same variable.

QTT is a system that resolves this conflict by recording usage of
variables with annotations from a semiring. It sits in the general
area of systems that use semiring annotatons to measure resource usage
\cite{BLAH}. The key feature of QTT, originally due to
\cite{mcbride16}, is that usage of variables in types counts for
$0$-usage in terms of the semiring used. This allows us to use normal
type theory as a specification language, while also enjoying the
benefits of linear typing for programs. The term typing judgement of
QTT has the following form:
\begin{displaymath}
  x_1 \stackrel{\rho_1}: S_1, \dots, x_n \stackrel{\rho_n}: S_n \vdash M \stackrel\sigma: T
\end{displaymath}
where the annotations $\rho_i$ are all from the semiring being
used. The annotation $\sigma$ is either $0$ or $1$, indicating whether
we are in the erased ($\sigma = 0$) fragment, where all the normal
rules of type theory apply, or the in the non-erased (``present'',
$\sigma = 1$) fragment, where a restricted typing discipline applies.
As we shall see below in the cases of $\Sigma$-types, iterable types,
and LFPL's $\Diamond$ type, the separation of QTT into these two
fragments allows an expressive combination of reasoning using full
type theory with the benefits of linear typing.

In the remainder of this sub-section, we describe the core of QTT. As
well as the term typing judgement given above, QTT also has judgements
for well-formed contexts ($\Gamma~\isctxt$) and types
($\Gamma \vdash T~\istype$), and definition equality of types and
terms ($\Gamma \vdash S \equiv T~\istype$ and
$\Gamma \vdash M \equiv N \stackrel\sigma: S$). It is an invariant of
the system that types are always well-formed in a context with all
annotations $0$, i.e., $\Gamma \vdash S~\istype$ implies
$0\Gamma = \Gamma$. An important admissible rule of the system, along
with weakening and substitution, is that of $0$-ing:
\begin{displaymath}
  \inferrule*
  {\Gamma \vdash M \stackrel1: S}
  {0\Gamma \vdash M \stackrel0: S}
\end{displaymath}
This rule allows us to take any term $M$ in the $\sigma = 1$ fragment
and treat it as if it were in the $\sigma = 0$ fragment, and hence use
it for specification purposes in types. As we add novel rules to QTT
in the following sections, we will be careful to maintain the
admissibility of this rule.

\subsubsection{Natural-number usages}

We use an instantiation of QTT with the natural number semiring, with
the usual semiring structure of addition and multiplication. In a mild
generalisation of the original presentation of QTT, we also allow
sub-usaging via the \emph{reverse} ordering on the naturals. That is,
if a variable is marked as usage $n$ and $m \geq n$, then we can also
regard it as usage $m$. This makes the system more like affine linear
logic, since $m \leq 0$ for all $m$, matching the system in
\autoref{sec:linear-intro}. We do \emph{not} have an unrestricted
usage $\omega$, since this would allow the possibility of unrestricted
duplication, and hence violate our polytime soundness properties.

\subsubsection{Contexts, variables, and conversion}

As we saw above, contexts in QTT are comprised of variable
$\stackrel\rho:$ type triples, where $\rho$ is a natural number
indicating how many times the variable $x$ is available for use in a
$\sigma = 1$ term. There are two operations on raw contexts: scaling
$\pi\Gamma$, which multiplies each $\rho$ in $\Gamma$ by $\pi$, and
addition $\Gamma_1 + \Gamma_2$, which adds two contexts' usage
annotations assuming that the lengths and types are equal.  The basic
usage-annotation discipline of QTT is demonstrated by the context
formation and variable rules:
\begin{mathpar}
  \inferrule*
  { }
  {\epsilon~\isctxt}

  \inferrule*
  {\Gamma~\isctxt \\ 0\Gamma \vdash S~\istype}
  {\Gamma, x \stackrel\rho: S~\isctxt}

  \inferrule*
  {0\Gamma, x \stackrel\sigma: S, 0\Gamma'~\isctxt}
  {0\Gamma, x \stackrel\sigma: S, 0\Gamma' \vdash x \stackrel\sigma: S}
\end{mathpar}
As with most dependent type theories, contexts are built inductively
from the empty context $\epsilon$ and extension of a context by a
variable with a type that is well-formed in the preceding
context. Usage annotations $\rho$ on variables are arbitrary, and
types are always judged in a $0$-annotated context. The variable rule
marks unused variables as usage $0$ and the selected variable with
usage $\sigma$.

\bob{Usage weakening somewhere}

% \begin{mathpar}

%   \inferrule* [right=Tm-Conv]
%   {\Gamma \vdash M \stackrel\sigma: S \\ 0\Gamma \vdash S \equiv T~\istype}
%   {\Gamma \vdash M \stackrel\sigma: T}
% \end{mathpar}

\subsubsection{$\Pi$- and $\Sigma$-types}

QTT's $\Pi$-types have the form $(x \stackrel\rho: S) \to T$,
indicating functions that, in the $\sigma = 1$ fragment, use their
arguments $\rho$-many times. The formation, introduction and
elimination rules are similar to the standard ones, except for the
addition of usage annotations:
\begin{mathpar}
  \inferrule*
  {0\Gamma \vdash S~\istype \\ 0\Gamma, x \stackrel0: S \vdash T~\istype}
  {0\Gamma \vdash (x \stackrel\rho: S) \to T~\istype}

  \inferrule*
  {\Gamma, x \stackrel{\sigma\rho}: S \vdash M \stackrel\sigma: T}
  {\Gamma \vdash \lambda x.M \stackrel\sigma: (x \stackrel\rho: S) \to T}

  \inferrule*
  {\Gamma_1 \vdash M \stackrel\sigma: (x \stackrel\rho: S) \to T \\
    \Gamma_2 \vdash N \stackrel{\sigma'}: S \\
    0\Gamma_1 = 0\Gamma_2 \\
    \sigma' = 0 \Leftrightarrow (\rho = 0 \lor \sigma = 0)}
  {\Gamma_1 + \rho\Gamma_2 \vdash M\,N \stackrel\sigma: T[N/x]}
\end{mathpar}
The side conditions on the elimination rule state that (i) both
$\Gamma_1$ and $\Gamma_2$ erase to the same context, so their sum is
defined; and (ii) the argument $N$ is erased (i.e., $\sigma' = 0$) iff
either the function does not use its argument, or we are in the
$\sigma = 0$ fragment and everything is being erased.

$\Sigma$-types are a little more involved, and demonstrate the
flexibility in QTT in allowing additional power in the $\sigma = 0$
fragment, where we do not need to care about
implementability. Formation and introduction are given by the rules:
\begin{mathpar}
  \inferrule*
  {0\Gamma \vdash S \\ 0\Gamma, x \stackrel0: S \vdash T}
  {0\Gamma \vdash (x \stackrel\pi: S) \otimes T}

  \inferrule*
  {\Gamma_1 \vdash M \stackrel{\sigma'}: S \\
    \Gamma_2 \vdash N \stackrel\sigma: T[M/x] \\\\
    0\Gamma_1 = 0\Gamma_2 \\
    \sigma' = 0 \Leftrightarrow (\pi = 0 \lor \sigma = 0)}
  {\pi\Gamma_1 + \Gamma_2 \vdash (M, N) \stackrel\sigma: (x \stackrel\pi: S) \otimes T}
\end{mathpar}
As with $\Pi$-typed, the first component of a $\Sigma$ type is
annotated with a usage for how many times it can be used, and this is
respected by the introduction rule. Elimination of $\Sigma$-types
depends on whether we are in the $\sigma = 0$ fragment or not. In the
$\sigma = 0$ fragment, we are free to disregard usage restrictions,
and use projections as normal:
\begin{mathpar}
  \inferrule*
  {\Gamma \vdash M \stackrel0: (x \stackrel\pi: S) \otimes T}
  {\Gamma \vdash \mathrm{fst}(M) \stackrel0: S}

  \inferrule*
  {\Gamma \vdash M \stackrel0: (x \stackrel\pi: S) \otimes T}
  {\Gamma \vdash \mathrm{snd}(M) \stackrel0: T[\mathrm{fst}(M)/x]}
\end{mathpar}
Hence, $\Sigma$-types act as normal in the $\sigma = 0$ fragment, and
we can use them as for normal type-theoretic constructions.  In the
$\sigma = 1$ fragment, we must take into account the resource content
of objects and use a pattern matching construct; the dependently typed
analogue of the $\otimes$-eliminator in \autoref{sec:lineartype}:
\begin{mathpar}
  \inferrule*
  {0\Gamma, z \stackrel0: (x \stackrel\pi: A) \otimes B \vdash C \\
    \Gamma_1 \vdash M \stackrel\sigma: (x \stackrel\pi: A) \otimes B \\
    \Gamma_2, x \stackrel{\sigma\pi}: A, y \stackrel{\sigma}: B \vdash N \stackrel\sigma: C[(x,y)/z]\\
    0\Gamma_1 = 0\Gamma_2}
  {\Gamma_1 + \Gamma_2 \vdash \mathrm{let}~(x,y) = M~\mathrm{in}~N \stackrel\sigma: C[M/z]}
\end{mathpar}

\subsubsection{The Identity Type}

QTT also supports an extensional equality type with equality reflection:
\begin{mathpar}
  \inferrule*
  {0\Gamma \vdash S~\istype \\
    0\Gamma \vdash M \stackrel0: S \\
    0\Gamma \vdash N \stackrel0: S}
  {0\Gamma \vdash M =_S N~\istype}

  \inferrule*
  {\Gamma \vdash M \stackrel\sigma: S}
  {\Gamma \vdash \conRefl(M) \stackrel\sigma: M =_S M}

  \inferrule*
  {\Gamma \vdash N \stackrel0: M_1 =_S M_2}
  {\Gamma \vdash M_1 \equiv M_2 \stackrel0: S}
\end{mathpar}
The equality type also has an $\eta$ rule demonstrating $\conRefl(M)$
as the only proof of equality \cite{hoffmann99dependent}. Note that
equality reflection only targets the $\sigma = 0$ fragment, we cannot
use equality reflection to convey realisability information.

\subsubsection{Universe} QTT can also be equipped in universe types,
just as in standard type theory. We refer to the rules presented in
\cite{atkey18qtt}.

\subsubsection{Datatypes} QTT, as we have presented it so far, has no
interesting base types to perform computation on. Following our
presentation of the simply typed linear systems in
\autoref{sec:linear-intro}, we add two kinds of datatype to QTT. First,
\autoref{sec:noniter-qtt}, we add non-iterable datatypes that allow
construction and case analysis, but no recursion. In
\autoref{sec:cons-free-qtt} we describe how to extend QTT to be a
dependently typed adaptation of the Cons-free system of
\autoref{sec:cons-free-intro}. In \autoref{sec:lfpl-qtt}, we apply the
same treatment to the LFPL-style system.

\subsection{Non-iterable Datatypes}
\label{sec:noniter-qtt}

\subsubsection{Booleans} The boolean type for QTT was described in
\cite{atkey18qtt}, but we repeat it here for completeness. Booleans
offer no possibility for iteration, but it is useful to see how the
QTT versions of the rules extend the simply typed rules from
\autoref{sec:linear-case-analysis} before moving to more complex
types.
\begin{mathpar}
  \inferrule*
  {\Gamma~\isctxt}
  {\Gamma \vdash \BoolTy~\istype}

  \inferrule*
  {\Gamma~\isctxt}
  {0\Gamma \vdash \cTrue, \cFalse \stackrel\sigma: \BoolTy}

  \inferrule*
  {0\Gamma_1, x \stackrel0: \BoolTy \vdash P~\istype \\
    \Gamma_1 \vdash M \stackrel\sigma: \BoolTy \\
    \Gamma_2 \vdash N_t \stackrel\sigma: P[\cTrue/x] \\
    \Gamma_2 \vdash N_f \stackrel\sigma: P[\cFalse/x] \\
    0\Gamma_1 = 0\Gamma_2}
  {\Gamma_1 + \Gamma_2 \vdash \If_{x.P}\: M \: \Then \: N_t \: \Else \: N_f \stackrel\sigma: P[M/x]}
\end{mathpar}
The introduction rules for booleans both use a $0$-d context,
indicating that construction of boolean values is free. Elimination of
booleans via a dependently typed if-then-else is more subtle with its
resource usage. The boolean to be eliminated must be constructed in a
context $\Gamma_1$, while the two branches are constructed in context
$\Gamma_2$. Since only one of the branches will be used, sharing
resources between the branches is expected.

One might wonder how, since constructing booleans is $0$-cost by their
introduction rules, the $\Gamma_1$ context will ever be non-$0$. This
is resolved by observing that booleans may be the output of processes
that consume time (e.g., the iteration constructs defined below), and
so $\Gamma_1$ will represent a requirement that the necessary resource
is provided.

\subsubsection{Lists} Lists are a little more complex than booleans,
because the $\cCons$ constructor takes two arguments, so their
resource usage must be combined. The type formation and introduction
rules are as follows:
\begin{mathpar}
  \inferrule*
  {0\Gamma \vdash T~\istype}
  {0\Gamma \vdash \ListTy(T)~\istype}

  \inferrule*
  {\Gamma \vdash T~\istype}
  {0\Gamma \vdash \cNil \stackrel\sigma: \ListTy(T)}

  \inferrule*
  {\Gamma_1 \vdash M \stackrel\sigma: T \\
    \Gamma_2 \vdash N \stackrel\sigma: \ListTy(T) \\
    0\Gamma_1 = 0\Gamma_2}
  {\Gamma_1 + \Gamma_2 \vdash \cCons(M,N) \stackrel\sigma: \ListTy(T)}
\end{mathpar}
Lists do have the potential for iteration by their recursive nature,
but in order to ensure the complexity guarantees we only permit
matching without recursion in the full calculus. Here is the rule for
dependently typed case analysis on lists:
\begin{mathpar}
  \inferrule*
  {0\Gamma_1, x \stackrel0: \ListTy(T) \vdash P~\istype \\
    \Gamma_1 \vdash M \stackrel\sigma: \ListTy(T) \\
    \Gamma_2 \vdash N_1 \stackrel\sigma: P[\cNil/x] \\
    \Gamma_2, h \stackrel\sigma: T, t \stackrel\sigma: \ListTy(T) \vdash N_2 \stackrel\sigma: P[\cCons(h,t)/x] \\
    0\Gamma_1 = 0\Gamma_2}
  {\Gamma_1 + \Gamma_2 \vdash \Match_{x.P}\,M\,\{\,\cNil \mapsto N_1; \cCons(h,t) \mapsto N_2\,\} : P[M/x]}
\end{mathpar}
In the $\sigma = 0$ fragment, however, we are free to iterate on lists
because computations in this fragment are only meant for type-level
computation, not for the program itself. Put another way, the type
checker may perform arbitary recursion on lists to type check the
program, but the program itself may not do so without correctly
accounting its costs as described in the following sections. The
$\sigma = 0$ recursor for lists has the following typing rule, which
is the standard dependent eliminator for lists, with everything
annotated as $0$ usage.
\begin{mathpar}
  \inferrule*
  {0\Gamma, x \stackrel0: \ListTy(T) \vdash P~\istype \\
    0\Gamma \vdash M \stackrel0: \ListTy(T) \\
    0\Gamma \vdash N_1 \stackrel0: P[\cNil/x] \\
    0\Gamma, h \stackrel0: T, t \stackrel0: \ListTy(T), p \stackrel0: P[t/x] \vdash N_2 \stackrel0: P[\cCons(h,t)/x]}
  {0\Gamma \vdash \mathrm{recList}_{x.P}\,M\,\{\,\cNil \mapsto N_1; \cCons(h,t;p) \mapsto N_2\,\} \stackrel0: P[M/x]}
\end{mathpar}

\subsection{Cons-free Natural Numbers and their Recursor}
\label{sec:cons-free-qtt}

The datatypes of the previous section still only allow us to write
programs in the $\sigma = 1$ fragment that are constant time in the
size of their input. As with the simply typed linear system, if we are
handed a list of an unknown length, we can only explore it to a fixed
depth determined statically. To write programs that do work
proportional to the size of their input, we need some form of iterable
datatype. In both our Cons-free and LFPL-style QTT systems, we use a
natural number datatype.

The cons-free system cannot allow the programmer to construct natural
numbers in the $\sigma = 1$ fragment, as this would violate the
complexity guarantees. However, we can use the flexibility of QTT to
allow free construction of naturals in the $\sigma = 0$ fragment,
which allows us to use natural numbers freely in types. Therefore, we
have the following introduction rules, only usable in the $\sigma = 0$
fragment:
\begin{mathpar}
  \inferrule*
  {\Gamma \vdash}
  {\Gamma \vdash \conZero \stackrel0: \tyNat}

  \inferrule*
  {\Gamma \vdash M \stackrel0: \tyNat}
  {\Gamma \vdash \conSucc(M) \stackrel0: \tyNat}
\end{mathpar}
The cons-free system allows free duplication of complete natural
numbers. This is accomplished by a special construct copying the
simply linear typed rule we gave above:
\begin{mathpar}
  \inferrule*
  {\Gamma \vdash M \stackrel\sigma: \tyNat}
  {\Gamma \vdash \mathrm{dupNat}(M) \stackrel\sigma: \tyNat \otimes \tyNat}
\end{mathpar}
Anyone who has reasoned about the metatheory of or implemented a type
checker for dependent types will view this rule with unease as it
appears to grant the ability to construct non-canonical values of pair
type, and consequently generate non-canonical naturals. We fix this by
adding an equational rule to the $\sigma = 0$ fragment, ensuring
\emph{definitionally} that $\mathrm{dupNat}$ acts as its name implies:
\begin{mathpar}
  \inferrule*
  {\Gamma \vdash M \stackrel0: \tyNat}
  {\Gamma \vdash \mathrm{dupNat}(M) \equiv (M,M) \stackrel0: \tyNat \otimes \tyNat}
\end{mathpar}
Note that this rule is well-typed by the $0$-needs-$0$ property of
QTT, and the fact that $0+0=0$.

The eliminator for these natural numbers takes the following
form. Disregarding the usage annotations, this has the same type
structure as the normal dependently typed recursor for natural
numbers:
\begin{displaymath}
  \mprset{flushleft}
  \inferrule*
  {0\Gamma, x \stackrel0: \tyNat \vdash P~\istype \\\\
    \Gamma \vdash M \stackrel\sigma: \tyNat \\\\
    0\Gamma \vdash N_z \stackrel\sigma: P[\conZero/x] \\\\
    0\Gamma, n \stackrel0: \tyNat, p \stackrel\sigma: P[n/x] \vdash N_s \stackrel\sigma: P[\conSucc(n)/x]}
  {\Gamma \vdash \tmRec_{x.P}\,M\,\{\conZero \mapsto N_z; \conSucc(n;p) \mapsto N_s\} \stackrel\sigma: P[M/x]}
\end{displaymath}
In the successor case, $N_s$, there are two bound variables: $n$ for
the natural number and $p$ for its induction hypothesis. Note that $n$
is required to be usage $0$ no matter what $\sigma$ is. We need the
variable $n$ to be present in order to correctly type the induction
hypothesis and the conclusion, but it must be marked as usage $0$ to
ensure that the resources captured by the number are not duplicated.

The reader is invited to compare this dependently typed rule with the
simply typed linear version in \autoref{sec:cons-free-intro}. Removing
the $0$-annotated parts of the rule, and the type dependency, yield
the exact same rule. Conversely, when $\sigma = 0$, this rule is
identical (up to $0$-annotations) to the usual dependently typed
recursor for natural numbers, and so we can use it in the types to
prove properties of programs just as we do in normal type theory. We
will see in \autoref{sec:cons-free-sound} that this rule is realisable
by polynomial-time amortised computation, and so is sound for
polynomial time.

\subsection{LFPL-style Diamonds and Natural Numbers}
\label{sec:lfpl-qtt}

As explained in \autoref{sec:lfpl-intro}, the LFPL system differs from
the Cons-free system in that it is possible to construct natural
numbers (and other iterable datatypes), provided one has the necessary
diamonds to pay for the construction. As with the natural number type
in the Cons-free system, it ought not be possible to construct
diamonds in the $\sigma = 1$ fragment, as this would amount to the
free distribution of diamonds to all which would lead to a collapse in
the complexity guarantees of the system. It is possible construct
diamonds in the $\sigma = 0$, though:
\begin{mathpar}
  \inferrule*
  {\Gamma \vdash}
  {0\Gamma \vdash \Diamond~\istype}

  \inferrule*
  {\Gamma \vdash}
  {0\Gamma \vdash * \stackrel0: \Diamond}

  \inferrule*
  {\Gamma \vdash M \stackrel0: \Diamond}
  {\Gamma \vdash M \equiv * \stackrel0: \Diamond}
\end{mathpar}
The $\Diamond$ type also supports an $\eta$-rule in the $\sigma = 0$
fragment, indicating that, in this fragment, it acts the same as a
unit type. This allows us to use freely use diamonds in types, and to
not have to care about the identity of particular diamonds, since by
this rule all diamonds are definitionally
equal\footnote{\emph{Fungible}, if one wishes to use a monetary
  metaphor.}.
% Link to Selinger et al: the $\eta$ rule for $\Diamond$s means that are
% essentially depending on the ``shape'' of anything that depends on
% diamonds.

Construction of natural numbers now requires a $\Diamond$ for
$\conZero$ and a $\Diamond$ and a predecessor for $\conSucc$:
\begin{mathpar}
  \inferrule*
  {\Gamma \vdash M \stackrel\sigma: \Diamond}
  {\Gamma \vdash \conZero(M) \stackrel\sigma: \tyNat}

  \inferrule*
  {\Gamma_1 \vdash M \stackrel\sigma: \Diamond \\
    \Gamma_2 \vdash N \stackrel\sigma: \tyNat \\
    0\Gamma_1 = 0\Gamma_2}
  {\Gamma_1 + \Gamma_2 \vdash \conSucc(M,N) \stackrel\sigma: \tyNat}
\end{mathpar}
In the $\sigma = 0$ fragment, we can construct $\Diamond$s for free,
and so construct natural numbers freely as well just as we did for the
Cons-free system above.

The dependently typed recursor for LFPL-style natural numbers again
augments the simply typed linear recursor from
\autoref{sec:lfpl-intro} with dependency information:
\begin{mathpar}
  \mprset{flushleft}
  \inferrule*
  {0\Gamma, x \stackrel0: \tyNat \vdash P~\istype \\\\
    \Gamma \vdash M \stackrel\sigma: \tyNat
    0\Gamma, d \stackrel1: \Diamond \vdash N_z \stackrel\sigma: P[\conZero(*)/x] \\\\
    0\Gamma, d \stackrel1: \Diamond, n \stackrel0: \tyNat, p \stackrel1: P[n/x] \vdash N_s \stackrel\sigma : P[\conSucc(*,n)/x]}
  {\Gamma \vdash \tmRec\,M\,\{\conZero(d) \mapsto N_z; \conSucc(d,n;p) \mapsto N_s\} \stackrel\sigma: P[M/x]}
\end{mathpar}
We have used $* : \Diamond$ as the value in the types for the zero and
successor cases. By the $\eta$-rule for diamonds, we could have
equally well used the $d$ variable that is in scope in each case.

Just as for the Cons-free system iterator above, in the $\sigma = 0$
fragment this rule is identical to the usual dependently typed
recursor for the natural numbers, so it can be used in the types to
reason about programs. Moreover, we will see in
\autoref{sec:lfpl-sound} that this rule is also sound for polynomial
time in a system with $\Diamond$s.

\subsection{Reflection of Realisability}
\label{sec:qtt-reflection}

Our final addition to QTT is \emph{reflection of realisability}. In
QTT thus far, it has been possible to reason about the non-resourced
behaviour of programs. This is because the $0$-ing process moving from
the $\sigma = 1$ fragment to the $\sigma = 0$ fragment erases all
resource information. This is sufficient for reasoning about the
extensional behaviour of programs, but it is useful to be able to make
statements like ``this function is realisable in polynomial time'' in
the types of QTT, something that is not current possible.

We remedy this by adding a \emph{realisable} type to QTT with the
following type formation and introduction and elimination rules:
\begin{mathpar}
  \inferrule*
  {0\Gamma \vdash A~\istype}
  {0\Gamma \vdash \Rtype(A)~\istype}

  \inferrule*
  {0\Gamma \vdash M \stackrel1: A}
  {0\Gamma \vdash \rIntro(M) \stackrel\sigma: \Rtype(A)}

  \inferrule*
  {\Gamma \vdash M \stackrel\sigma: \Rtype(A)}
  {\Gamma \vdash \rElim(M) \stackrel{\sigma'}: A}
\end{mathpar}
Intuitively, the type $\Rtype(A)$ is inhabited whenever the type $A$
is realisable in the $\sigma = 1$ fragment of the system. In
particular, the type $\Rtype(\tyNat \to \tyNat)$ is the type of all
realisable functions from natural numbers to natural numbers. In the
polynomial time systems we are concerned with here, this is exactly
the type of polynomial functions. Note that in the introduction rule,
the premise is required to be in the $\sigma = 1$ fragment, to ensure
that the type is realisable, while in the elimination rule, the
conclusion is in an arbitrary fragment $\sigma'$. This flexibility is
require to maintain the admissibility of the $0$-ing rule.

With just the rules given here, the type $\Rtype(A)$ is nothing more
than a statement that a given type is realisable with a polytime
implementation. This is enough to do some proofs that we present in
the next section, e.g., that polytime functions are closed under
composition, but one could imagine stronger reflection principles that
allow deeper logical consequences of polytime realisability to be
proved internally. We discuss this further in
\autoref{sec:towards-synthetic-complexity-theory}.

Readers familiar with Benton's Linear/Non-linear system
\cite{benton94} will note that the $\Rtype(A)$ constructor is the QTT
analogue of the right adjont $G$ type constructor in that system. The
$\Sigma$-types describe above play the role of the left adjoint $F$
types, in a similar way to the dependent linear type system of Pradic,
Krishnaswami and Benton \cite{pradic}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Programming and Proving with Polytime}
\label{sec:programming-polytime}

% \begin{enumerate}
% \item Non-deterministic polytime via a bit oracle -- needs a statement
%   of success
% \item Previous could also be interpreted as probabilistic polytime
%   (access to coin flips); can we formalise some kind of cryptographic
%   property this way? Have a look at hash functions in the
%   Authenticated Data Structures paper.
% \item Insertion sort, quicksort as polytime functions.
% \item A type of polytime functions; via a reflection operator
% \item Programming with explicit resource annotations, including
%   reasoning about the resource requirements directly.
% \end{enumerate}

\subsection{Building Datatypes}
\label{sec:prog-datatypes}

We have only defined an iterable natural number datatype for both of
our systems above. Using the power of dependent types with a universe
we can use these natural numbers to create further datatypes whose
size is measured by a natural. Iteration on the size yields iteration
over the full datastructure. For example, in the LFPL system, we can
define a type of iterable lists by pairing a size with a type of
elements defined by recursion:
\begin{displaymath}
  \mathrm{IList}(A) = \Sigma n \stackrel1: \tyNat.\,\mathrm{rec}_{x.\mathsf{U}}\,n\,\{ \conZero(d) \mapsto I; \conSucc(d,n;p) \mapsto A \otimes p \}
\end{displaymath}

Nil and Cons constructors.

List iterator:

\bob{U notation for the universe above}

\begin{enumerate}
\item Insertion sort in LFPL.
\item Insertion sort in LFPL with correctness specification.
\end{enumerate}

\subsection{Polytime Problems}

Uses

\begin{displaymath}
  (A : \mathsf{U}, P : A \to \mathsf{U})
\end{displaymath}

Cartesian container morphisms

\subsection{Polytime-based Complexity Classes}

Monads

\subsubsection{Non-deterministic Polynomial Time}

\subsubsection{Probabilistic Polynomial Time}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Polytime Soundness via Realisability}
\label{sec:soundness}

\newcommand{\cstaccess}{1}
\newcommand{\cstmkclo}{1}
\newcommand{\cstapp}{1}
\newcommand{\cstmkpair}{1}
\newcommand{\cstmkunit}{1}
\newcommand{\cstTrue}{1}
\newcommand{\cstFalse}{1}
\newcommand{\cstLetpair}{1}
\newcommand{\cstSeq}{1}
\newcommand{\cstIf}{1}

% \newcommand{\cstaccess}{c_{\mathit{access}}}
% \newcommand{\cstmkclo}{c_{\mathit{mkclo}}}
% \newcommand{\cstapp}{c_{\mathit{app}}}
% \newcommand{\cstmkpair}{c_{\mathit{mkpair}}}
% \newcommand{\cstmkunit}{c_{\mathit{mkunit}}}
% \newcommand{\cstTrue}{c_{\mathit{mktrue}}}
% \newcommand{\cstFalse}{c_{\mathit{mkfalse}}}
% \newcommand{\cstLetpair}{c_{\mathit{letpair}}}
% \newcommand{\cstSeq}{c_{\mathit{seq}}}
% \newcommand{\cstIf}{c_{\mathit{if}}}

\newcommand{\clo}[2]{\mathsf{clo}\langle #1 , #2 \rangle}
\newcommand{\synTrue}{\mathsf{true}}
\newcommand{\synFalse}{\mathsf{false}}

\newcommand{\ExpSet}{\mathcal{E}}
\newcommand{\ValSet}{\mathcal{V}}

\newcommand{\rplus}{\oplus}
\newcommand{\rzero}{\emptyset}

In this section and the next, we establish the polytime soundness of
our extensions of QTT by adapting a realisability method due to
\cite{hofmann-dallago}. This approach is based on a three way coupling
between abstract mathematical elements (the \emph{what}), values from
a machine model (the \emph{how}), and resource potentials (the
\emph{fuel}). Each type in the system is defined as a three way
relation between these elements. The set of abstract elements depends
on the type being interpreted (e.g., types of natural numbers will be
defined in terms of the set $\mathbb{N}$). The machine model is fixed
across all types. We describe the particular machine model we use for
this paper in \autoref{sec:machine-model}. Potentials are arranged
into \emph{resource monoids} that we define in
\autoref{sec:resource-monoids}. Following Hofmann and Dal Lago, we
will select resource monoids appropriate to the kind of system that we
want to prove complexity soundness for.

\paragraph{Agda Formalisation} The key soundness results in this
section have been formalised in the Agda proof assistant
\cite{agda}. The Agda formalisation can be found in the supplementary
material. After each definition and result we provide a pointer to the
Agda modules where the corresponding formalisation can be found, and
note interesting features of the mechanisation.

\subsection{Machine Model and Operational Semantics}
\label{sec:machine-model}

\begin{figure}
  \centering
  {\bf Syntax}
  \begin{displaymath}
    \begin{array}{lcl}
      i,j &\in&\mathbb{N} \\
      E \in \ExpSet &::=& \lambda E \mid * \mid (i, j) \mid \synTrue \mid \synFalse \mid i \mid \Let\,E_1\,\In\,E_2 \mid i \cdot j \mid \LetPair\,i\,\In\,E \mid \If\,i\,E_1\,E_2 \\
      V \in \ValSet &::=& \clo{E}{\eta} \mid * \mid (V_1, V_2) \mid \synTrue \mid \synFalse \\
      \eta &::= & [] \mid \eta :: V
    \end{array}
  \end{displaymath}

  \vspace{1em}

  {\bf Evaluation: Construction}
  \begin{mathpar}
    \inferrule* [right=MkClo]
    { }
    {\lambda E , \eta \Downarrow_{\cstmkclo} \clo{E}{\eta}}

    \inferrule* [right=MkUnit]
    { }
    {*, \eta \Downarrow_{\cstmkunit} *}

    \inferrule* [right=MkPair]
    {\eta[i] = V_1 \\ \eta[j] = V_2}
    {(i, j), \eta \Downarrow_{\cstmkpair} (V_1, V_2)}

    \inferrule* [right=MkTrue]
    { }
    {\synTrue, \eta \Downarrow_{\cstTrue} \synTrue}

    \inferrule* [right=MkFalse]
    { }
    {\synFalse, \eta \Downarrow_{\cstFalse} \synFalse}
  \end{mathpar}

  \vspace{1em}

  {\bf Evaluation: Variable access and Sequencing}
  \begin{mathpar}
    \inferrule* [right=Access]
    {\eta[i] = v}
    {i, \eta \Downarrow_{\cstaccess} v}

    \inferrule* [right=Seq]
    {E_1, \eta \Downarrow_{k_1} V \\
      E_2, (\eta :: V) \Downarrow_{k_2} V'}
    {\Let\,E_1\,\In\,E_2, \eta \Downarrow_{k_1 + \cstSeq + k_2} V'}
  \end{mathpar}

  \vspace{1em}

  {\bf Evaluation: Elimination}
  \begin{mathpar}
    \inferrule* [right=App]
    {\eta[i] = \clo{E}{\eta'} \\
      \eta[j] = V \\
      E , (\eta' :: \clo{E}{\eta'} :: V) \Downarrow_k V'}
    {(i \cdot j), \eta \Downarrow_{\cstapp + k} V'}

    \inferrule* [right=LetPair]
    {\eta[i] = (V_1, V_2) \\
      E, (\eta :: V_1 :: V_2) \Downarrow_k V}
    {\LetPair\,i\,\In\,E, \eta \Downarrow_{\cstLetpair + k} V}

    \inferrule* [right=IfTrue]
    {\eta[i] = \synTrue \\ E_1, \eta \Downarrow_k V}
    {\If\,i\,E_1\,E_2, \eta \Downarrow_{\cstIf + k} V}

    \inferrule* [right=IfFalse]
    {\eta[i] = \synFalse \\ E_2, \eta \Downarrow_k V}
    {\If\,i\,E_1\,E_2, \eta \Downarrow_{\cstIf + k} V}
\end{mathpar}
  \caption{Language with CBV Big-step Costed Evaluation Semantics}
  \label{fig:opsem}
\end{figure}

We demonstrate that every program that can be written in extensions of
QTT has the complexity bounds that we claim by translating QTT terms
into an untyped CBV $\lambda$-calculus with a costed operational
semantics. The syntax and rules of our target language are given in
\autoref{fig:opsem}.

Variables are represented as de Bruijn indicies $i, j$. Expressions
$E \in \ExpSet$ can be (anonymous) $\lambda$-abstractions, unit,
pairing and boolean values, variables, sequencing, application, pair
elimination, and conditionals. Note that, with the exception of
$\lambda$-abstraction and sequencing, expressions never contain nested
expressions; instead referring to variables already defined. Values
$V \in \ValSet$ can be closures $\clo{E}{\eta}$, where $\eta$ is an
environment for the closure, unit values, pairs and booleans.

Costed evaluation of expressions in environments is defined by a
big-step operational semantics $E, \eta \Downarrow_k V$, where $k$ is
the number of steps. For simplicity, all operations cost $1$ unit,
though this could be generalised to allow for different operations to
have different costs. We use $\eta[i]$ to access the $i$th variable in
the environment, counting from the right. The evaluation rules are
mostly as one would expect, except for the application rule which
includes a self reference to the closure being invoked, in order to
allow recursive definitions.

\paragraph{Agda Formalisation} The machine model is defined in the
Agda module \AgdaModule{MachineModel}. We use an intrinsicaly
well-scoped syntax, which ensures that all variable look up operations
are always well defined.

\subsection{Resource Monoids}
\label{sec:resource-monoids}

As we mentioned above, resource potentials are attached to values to
represent the amount of intrinsic potential they have to fuel
computation. Resource potentials are organised into resource
monoids. To be able to account for the combined potential attached to
composite data and programs (e.g., pairs, or functions applied to
arguments) we will require monoid structure on potentials. The action
of turning potential difference into fuel for computation will be
modelled by a difference function. Finally, we require that our
resource monoid contains sufficient elements to fuel constant time
operations. We gather these requirements into a formal definition as
follows, which is a reformulation of the resource monoids of
\cite{hoffman-dallago}:

\begin{definition}
  A \emph{resource monoid} $M$ consists of:
  \begin{enumerate}
  \item A carrier set $|M|$, whose elements represent amounts of
    potential. We use Greek letters $\alpha$, $\beta$, $\gamma$ to
    denote elements of a resource monoid.
  \item Commutative monoid structure $(\rplus,\rzero)$ on $|M|$, so
    we can add potentials.
  \item a \emph{difference function} $M : |M| \times |M| \to \natinf$,
    where $\natinf$ is the natural numbers extended with a negative
    infinity $- \infty$ and $- \infty + k = -\infty$. A difference
    $M(\alpha, \beta) = k \in \mathbb{N}$ means that starting with
    potential $\alpha$ and ending with potential $\beta$ yields $k$
    units of fuel. A difference of $- \infty$ means that $\alpha$
    contains insufficient potential to reach $\beta$. Differencing
    must satisfy:
    \begin{enumerate}
    \item for all $\alpha$, $M(\alpha,\alpha) = 0$; and
    \item for all $\alpha, \beta, \gamma$,
      $M(\alpha, \beta) + M(\beta, \gamma) \leq M(\alpha, \gamma)$.
    \end{enumerate}
    The latter is a ``reverse triangle inequality'': the fuel
    recoverable by moving between potential levels $\alpha$ and
    $\gamma$ via $\beta$ may be less than the fuel recoverable
    moving from $\alpha$ to $\gamma$ directly.
  \item Differencing and the commutative monoid structure must satisfy:
    \begin{enumerate}
    \item $M(\alpha, \beta) \leq M(\alpha \rplus \gamma, \beta \rplus \gamma)$; and
    \item $M(\alpha, \rzero) = 0$.
    \end{enumerate}
  \item An \emph{accounting function}
    $\mathit{acct} : \mathbb{N} \to |M|$ such that for all $k$,
    $k \leq M(\mathit{acct}(k),\rzero)$.
  \end{enumerate}
\end{definition}

For any resource monoid $M$, we can define an action of $\mathbb{N}$
on $M$ as $n \cdot \alpha = \alpha \oplus \cdots \oplus \alpha$, where
the right-hand side has $n$ summands.

\paragraph{Alternative definition} Every resource monoid induces a
partial ordering on its carrier set by $\alpha \leq \beta$ iff
$0 \leq M(\alpha, \beta)$. Taking this idea further, we can
reformulate a resource monoid as a symmetric monoidal category
enriched in the symmetric monoidal category $\natinf$, where the
monoid structure is addition. Isomorphism of objects $\alpha, \beta$
in such a category amounts to having $M(\alpha, \beta) = 0$. The
conditions in the definition above amount to the usual identity and
composition laws for enriched categories.

\paragraph{Agda Formalisation} Resource monoids are defined in the
module \AgdaModule{Algebra.ResourceMonoid}. We use a formulation
closer to the enriched category theory definition for the actual
formalisation, because it avoids having to treat equality in the
monoid structure separately from the induced preorder on elements.

\subsubsection{Specific Resource Monoids}

The simplest example of a resource monoid is given by the natural
numbers $\mathbb{N}$, where each number stands directly an amount of
stored fuel.

\begin{definition}[Natural Number Resource Monoid]
  Monoid structure is given by normal addition. Differencing is
  defined as
  \begin{displaymath}
    \mathbb{N}(m,n) = \left\{
      \begin{array}{ll}
        m - n & m \geq n \\
        - \infty & \textrm{otherwise}
      \end{array}\right.
  \end{displaymath}
  and $\mathit{acct}(k) = k$. Note that this is the simplest possible
  resource monoid due to the requirement that the $\mathit{acct}$
  function must exist.
\end{definition}

\newcommand{\MaxPoly}{\mathrm{MaxPoly}}
\newcommand{\PlusPoly}{\mathrm{PlusPoly}}

The differencing operator of the natural number resource monoid can
only supply as much fuel as is contained in the potential.
% This makes
% it suitable for amortised analysis, where we directly store potential
% in data structures.
For the two polynomial time systems, we need more sophisticated
structures, both originally presented by Dal Lago and Hofmann. The
fundamental idea with both is to represent potentials as pairs
$(m,p)$, where $m$ is a natural number and $p$ is a polynomial. The
$m$ tracks the ``size'' of data as it pertains to the number of times
an operation will be repeated by iterating over it --- for example, an
iterable natural number will have size equal to itself, but a
non-iterable natural number may be assigned zero size. The polynomial
$p$ tracks the complexity of a program as a function of the size of
the input. This leads to a differencing operator that evaluates the
polynomial with the size of the data:

\begin{definition}[Polynomial Resource Monoids]\label{def:polynomial-resource-monoids}
  The \emph{Max-Polynomial} resource monoid $\MaxPoly$ has carrier set
  consisting of pairs $(m,p)$ where $m$ is natural number and $p$ is a
  polynomial with natural number coefficients. Addition of elements is
  defined as $(m,p) \oplus (n,q) = (m \sqcup n, p+q)$, where $\sqcup$
  is the max operator, with $\emptyset = (0,0)$. Difference is defined
  as:
  \begin{displaymath}
    \MaxPoly((m,p),(n,q)) = \left\{
      \begin{array}{ll}
        p(m) - q(m)&m \geq n \textrm{ and } \forall k \geq m. p(k) \geq q(k) \\
        - \infty & \textrm{otherwise}
      \end{array}
    \right.
  \end{displaymath}
  $\MaxPoly$ accounts for constant time with constant polynomials:
  $\mathit{acct}(k) = (0,\lambda x.k)$.

  The \emph{Plus-Polynomial} resource monoid $\PlusPoly$ is defined
  the same way as $\MaxPoly$ except that the monoid addition adds the
  natural number components instead of taking their maximum:
  $(m,p) \oplus (n,q) = (m + n, p+q)$.
\end{definition}

It is perhaps easier to see how the differencing operator works in the
special case of the difference $\MaxPoly((m,p),(0,0)) = p(m)$. I.e.,
if we have code that contains data of size $m$ and a program with
complexity $p$, then running the combination with no expectation of
remaining potential yields $p(m)$ available steps. The $\MaxPoly$ and
$\PlusPoly$ resource monoids will be used for the \emph{Cons-free} and
LFPL-style systems respectively, as we explain in
\autoref{sec:realising-iteration} and show how these resource monoids
yield the required polytime bounds on programs.

\paragraph{Agda Formalisation} The natural number resource monoid is
defined in the Agda module \AgdaModule{Algebra.ResourceMonoid.Nat} and the
 polynomial monoids are both defined in
\AgdaModule{Algebra.ResourceMonoid.Polynomial}. The definition is
parameterised by the ``size monoid'' operation (either $\sqcup$ or
$+$) used to compose sizes.

% FIXME: difference between this definition and Hofmann-Dal Lago's: we
% use polynomials with natural number coefficients instead of a function
% that is bounded by a polynomial.

\subsubsection{Resource sub-monoids}

The separation between sizes of data and complexity of code in the
polynomial resource monoids motivates the use of resource sub-monoids
to ensure that programs themselves (as opposed to higher order code
which may contain closed over data) do not contain data that can be
iterated. We do this by requiring that programs' potential must come
from a specified resource sub-monoid:

\begin{definition}[Resource Sub-Monoids]
  A \emph{resource sub-monoid} $M_0 \subseteq M$ of a resource monoid
  $M$ consists of a subset $|M_0| \subseteq |M|$ that is closed under
  the monoid operations and $\mathit{acct}$.
\end{definition}

For both $\MaxPoly$ and $\PlusPoly$, the elements with zero size
component, i.e., of the form $(0,p)$, form a resource sub-monoid that
we will use for interpreting programs. We will call these sub-monoids
$\MaxPoly_0$ and $\PlusPoly_0$.

\subsection{Models of Quantitative Type Theory from Indexed Preorders}
\label{sec:qtt-models}

\cite{atkey18qtt} described a general class of QTT models termed
\emph{Quantitative Categories with Families} (QCwFs). This class of
models fits well with the syntax of QTT, but to actually build models
it is convenient to construct them from certain indexed linear
preorders. The BCI-realisability model construction in
\cite{atkey18qtt} is a special case of this construction, but most of
the details are the same, so we only sketch a special case here to
concentrate on the specifics of complexity bounded realisability.

\begin{definition}\label{def:indexed-linear-preorder}
  A \emph{$\mathbb{N}$-linear preorder} is a preordered set
  $(L, \leq)$ with
  \begin{enumerate}
  \item a commutative monoid $(I, -\otimes-)$ that is monotone
    w.r.t. the order;
  \item is closed: there is an operation
    $\multimap : L \times L \to L$ such that $x \otimes y \leq z$ iff
    $x \leq y \multimap z$; and
  \item has a function $\oc : \mathbb{N} \to L \to L$, to interpret
    resource requirement adjustments, which satisfies the following
    inequalities:
    \begin{enumerate}
    \item $\oc_0 X \simeq I$, for discarding;
    \item $\oc_{m+n} X \leq (\oc_m X) \otimes (\oc_n X)$, for duplication;
    \item $\oc_m\oc_n X \leq \oc_{mn} X$ for nesting;
    \item $\oc_1 X \leq X$ for extraction / dereliction;
    \item $(\oc_n X) \otimes (\oc_n Y) \leq \oc_n(X \otimes Y)$, for distribution; and
    \item $n \leq m$ implies $\oc_n X \leq \oc_m X$, for usage weakening.
    \end{enumerate}
  \end{enumerate}
  The collection of all linear preorders and functions that preserve the
  order and the operations forms a category $\LinPreorder$.

  An \emph{indexed linear preorder} $L : \Set^\op \to \LinPreorder$ is
  a contravariant function, where we write $f^* : L(B) \to L(A)$ for
  the action of $L$ on functions $f : A \to B$, such that such that
  reindexing along projections has a right adjoint
  $L_{\Sigma_{a \in A}.B}(\pi_1^*X, Y) \cong L_{A}(X, \forall_B Y)$
  that commutes with reindexing.
\end{definition}

Given an indexed linear preorder $L : \Set^\op \to \LinPreorder$, we
construct a QCwF model of QTT with the basic type formers from
\autoref{sec:qtt-basic}:
\begin{enumerate}
\item Define a category $\cat{L}$ of interpretations of contexts with
  objects that are pairs $(A \in \Set, X \in L(A))$ and morphisms
  $f : (A,X) \to (B, Y)$ that are functions $f : A \to B$ such that
  $X \leq f^*Y$ (this is the Grothendieck category of $L$). There is a
  faithful functor $U : \cat{L} \to \Set$. The category $\cat{L}$ will
  be used for interpreting contexts in the $\sigma = 1$ fragment of
  QTT.
\item Define scaling of objects of $\cat{L}$ by
  $\pi(A,X) = (A, \oc_\pi X)$, and addition of $(A, X)$ and $(A, Y)$
  as $(A, X \otimes Y)$.
\item For each set $A$, define the collection of semantics types
  $\Ty(A)$ as the collection of $B : A \to \Set$ and
  $X \in L(\Sigma_{a \in A}.B(a))$. Thus a QTT type consists of an
  extensional meaning $B$ and its realisability specification $X$.
\item For each $A$ and $(B,X) \in \Ty(A)$, the $\sigma = 0$ fragment
  terms $\Tm(A,(B,X))$ are functions $\Pi_{a \in A}.\,B(a)$. For each
  context interpretation $(A, X)$ in $\cat{L}$ and type intepretation
  $(B,Y) \in \Ty(A)$, the $\sigma = 1$ fragment terms
  $\RTm((A,X),(B,Y))$ are functions $f : \Pi_{a \in A}.\,B(a)$ such
  that $X \leq \overline{f}^*Y$, where
  $\overline{f} : A \to \Sigma_{a \in A}.\,B(a)$ is the section
  associated with $f$.
\item The empty context is interpreted as $(\{*\}, I)$ and context
  extension $(A,X).(B,Y)$ (i.e., comprehension) by
  $(\Sigma_{a \in A}.B(a), \pi_1^* X \otimes Y)$.
\item Given $(A,X) \in \Ty(C)$ and
  $(B,Y) \in \Ty(\Sigma_{c \in C}.A(c))$, $\Sigma$-types are
  interpreted similarly to context extension and $\Pi$-types are
  interpreted as
  $(\lambda c.~(\Pi_{a \in A(c)}.\,B(c,a)), \forall_{A}(X \multimap
  (\mathit{ev}\,f)^*Y))$, where
  $\mathit{ev}\,f : (\Sigma_{c \in C}.A(c)) \to (\Sigma_{c \in
    C}.\Sigma_{a \in A(c)}.B(c,a))$ is defined using application of
  $f$.
\item Universe and Equality types are interpreted as normal in $\Set$
  with the realisability component set to $I$ in both cases. Note that
  the universe of small types includes realisability information for
  each type.
\item Realisability reflection for a type $(B,X) \in \Ty(A)$ is
  intepreted as the type
  $(\lambda a. \{ b \in B(a) \mid I \leq (\lambda a. (a,b))^*X \},
  I)$. Thus the set-component of the type is restricted to the
  elements that are realisable, while the actual realisability
  component is the ``empty'' $I$ realisability specification.
\end{enumerate}

\paragraph{Agda Formalisation} The indexed linear preorders are
defined in the Agda module \AgdaModule{IndexedLinear}. We have not yet
completed a formalisation of the construction of a full model of QTT
from an indexed linear preorder, nor have we constructed the syntax of
QTT in Agda, so these parts are currently unmechanised.

\subsection{Amortised Complexity Realisability Model}
\label{sec:realisability-model}

Equipped with our underlying costed model of computation
(\autoref{sec:machine-model}) and a compositional notion of resource
potential (\autoref{sec:resource-monoids}), we can construct models of
QTT that witness the resource and type soundness of our complexity
constrained systems. We fix a resource monoid $M$ with distinguished
sub-monoid $M_0$.

\subsubsection{Indexed Linear Preorder}
\label{sec:realisability-indexed-linear-preorder}

We now define an indexed linear poset $L$ of realisers over $\Set$
that ties together our ``mathematical'' model of types in $\Set$ with
our machine model and resource monoid. This construction is a
reformulation of \cite{hoffmann-dallago} suitable for modelling
dependent types. For a set $A$, the carrier of $L(A)$ is the set of
ternary relations $X \subseteq A \times M \times \ValSet$ and we
define the ordering $X \leq Y$ to hold iff there exists realising
expression $E \in \ExpSet$ and potential $\gamma \in M_0$ such that
for all $a \in A$, $\alpha \in M$ and $v \in \ValSet$ with
$(a,\alpha,v) \in X$, we have that there exists a result
$v' \in \ValSet$, step count $k \in \mathbb{N}$ and result potential
$\beta \in M$ with:
\begin{enumerate}
\item $E, v \Downarrow_k v'$ (evaluation successfully completes in $k$ steps);
\item $(a, \beta, v') \in Y$ (the result is well-resourced and
  satisfies $Y$); and
\item $k \leq M(\alpha \rplus \gamma, \beta)$ (the step count is within the
  difference between the initial potential and the result potential).
\end{enumerate}
Note that the definition of realisablity is uniform in the element $a$
-- the realising expression $e$ and the potential $\gamma$ must work
for all $a$ -- thus the implementation and complexity measure of the
transition being modelled cannot depend on what the input is. Put in
implementation terms, the input $a$ is not present at
runtime. Moreover note that the potential $\gamma$ attached to the
expression $e$ must come from the sub-monoid $M_0$, indicating that is
intended to be data-free, while the potential $\alpha$ for the input
is from the full monoid $M$, so it can contain data and functions.

For $X, Y \in L(A)$, the required elements for symmetric monoidal
closed structure are defined as follows. For the tensor product
$X \otimes Y \in L(A)$, the realising value must be a pair $(v_1,v_2)$
and the potential of the pair must split into suitable potentials
$\alpha_1$, $\alpha_2$ for the components. For the residual
$X \multimap Y$, the realising value must be a closure with potential
to, when added to the potential of an input, compute the output with
enough remaining. Note that the potential attached to a closure
($\alpha$, here) need not be from the sub-monoid $M_0$. Unlike
top-level term interpretations, closures may contain data.
\begin{displaymath}
  \begin{array}{lcl}
    X \otimes Y &=& \{ (a, \alpha, (v_1, v_2)) \mid \exists \alpha_1, \alpha_2.~0 \leq M(\alpha, \alpha_1 \rplus \alpha_2) \land X(a,\alpha_1,v_1) \land Y(a,\alpha_2,v_2) \}\\
    X \multimap Y &=& \{ (a, \alpha, \clo{E}{\eta}) \mid
                      \begin{array}[t]{@{}l}
                        \forall \alpha' \in M, v,w \in \ValSet.\,X(a,\alpha',v) \Rightarrow\\
                        \quad \exists v', k, \beta.\,
                        E, (\eta :: w :: v) \Downarrow_k v' \land Y(a,\beta,v') \land k \leq M(\alpha \rplus \alpha', \beta) \} \end{array}
  \end{array}
\end{displaymath}
The seemingly useless $w \in \ValSet$ in the formula for
$X \multimap Y$ is a dummy argument standing for the self-referential
reference to the closure used for defining recursive programs.

Each $L(A)$ has a terminal (i.e. top) element, which is also the unit
for $\otimes$, defined as
$I_A = \{(a, \alpha, *) \mid a \in A, \alpha \in M\}$. The potential
$\alpha$ here is unrestricted, so $I_A$ can consume an arbitrary
resource.

$\mathbb{N}$-Graded exponentials in each $L(A)$ are defined using the
action $(\cdot)$ of $(\mathbb{N}, \leq)$ on $M$ defined above. At the
level of the realising values, the modality $\oc_n$ has no effect; it
only serves to alter the required potentials.
\begin{displaymath}
  \oc_n\,X = \{(a,\alpha,v) \mid \exists \alpha'.\,M(n \cdot \alpha', \alpha) = 0 \land (a, \alpha', v) \in X \}
\end{displaymath}

$L$ also has arbitrary $\Set$-indexed products, realised ``lazily'' as
functions that take dummy arguments. For $A \in \Set$ and
$B \in A \to \Set$ and $X \in L(\Sigma A.\,B)$, we define
$\forall_B X \in L(A)$ similarly to $\multimap$ above, but with
different resource and indexing requirements:
\begin{displaymath}
  \forall_B X = \{ (a,\alpha,\clo{E}{\eta}) \mid \forall b, v.~\exists v', \beta, k. E, (\eta :: v :: *) \Downarrow_k v' \land X((a,b),\beta,v') \land k \leq M(\alpha,\beta) \}
\end{displaymath}
Note, as with the definition of $X \leq Y$ above, the realiser closure
$\clo{E}{\eta}$ must be chosen uniformly for all $b$. This definition
also appears to allow arbitrary computation (paid for by $\alpha$) to
happen when the realising closure is applied, but the potential
$\alpha$ will only ever be greater than $\beta$ by enough to handle
the administrative costs of applying the function.

To complete the construction of $L$ as an indexed linear preorder, we
need to give realisers for each of the required inequalities in
\defref{def:indexed-linear-preorder}. In each case, this is a matter
of programming in the language of \autoref{sec:machine-model}. For
example, transitivity of the order is realised by sequencing of
realising expressions. The corresponding potentials are calculated by
counting the steps in the ensuing programs.

\begin{proposition}\label{prop:realisability-model}
  $L$, with $I$, $\otimes$, $\multimap$, $\oc_n$, and $\forall_{B}$
  defined above, is an indexed linear preorder.
\end{proposition}

\paragraph{Agda Formalisation} The construction of this indexed linear
preorder and the proof of \propref{prop:realisability-model} are
formalised in the Adga module
\AgdaModule{AmortisedRealisabilityModel}.

\subsubsection{Non-iterable Data Types}
\label{sec:noniter-model}

The model of QTT constructed in \propref{prop:realisability-model}
does not yet include any useful base types. Handling iterable types,
which are the ones that induce non-constant-time complexities,
requires specific properties of resource monoids that we introduce in
\autoref{sec:realising-iteration}.

Rather than perform a complex generic construction of datatypes to
show how non-iterable versions may be realised, we focus on the two
representative examples from \autoref{sec:linear-case-analysis} and
\autoref{sec:noniter-qtt}. Booleans are the simplest case, with only
two cases and no chance of iteration. Lists are more complex; we can
have non-iterable lists containing iterable data.

\paragraph{Booleans} Fix $\mathbb{B} = \{ \mathit{tt}, \mathit{ff} \}$
as our set of boolean elements. We define an element of
$L(\mathbb{B})$ to represent boolean values:
\begin{displaymath}
  \mathrm{Bool} = \{ (\mathit{tt}, \alpha, \synTrue) \mid \alpha \in M \} \cup \{ (\mathit{ff}, \alpha, \synFalse) \mid \alpha \in M \}
\end{displaymath}
Thus, the boolean $\mathit{tt}$ is represented by the value $\synTrue$
and $\mathit{ff}$ is represented by $\synFalse$. In both case, we
allow arbitrary potentials to be attached.

Realisability of the construction and elimination of booleans amounts
to the existence of the following inequalities. In an arbitrary fibre
$L(A)$, we have $I_A \leq \mathit{tt}^* \mathrm{Bool}$ and
$I_A \leq \mathit{ff}^* \mathrm{Bool}$ (treating $\mathit{tt}$ and
$\mathit{ff}$ as constant functions $A \to \mathbb{B}$). These
inequalities are realised by the corresponding $\synTrue/\synFalse$
expression. For conditionals, the types involved are a little more
complex to ensure agreement between boolean manipulations at the
$\Set$-level and the realising computations. To get a realiser for a
condition, we require a set $A$, an element $X \in L(A)$ (standing for
the context) and an element $Y \in L(A \times \mathbb{B})$ (standing
for the target type) and the existence in $L(A)$ of inequalities
$X \leq (\lambda a. (a, \mathit{tt}))^* Y$, for the true case, and
$X \leq (\lambda a. (a, \mathit{ff}))^* Y$, for the false case. When
we have all these, we get in $L(A \times \mathbb{B})$ an inequality
$\pi_1^* X \otimes \pi_2^* \mathrm{Bool} \leq Y$. This construction
suffices to realise the rules for QTT booleans in
\autoref{sec:noniter-qtt}.

\paragraph{Lists} Lists are a little more involved, due to the need to
explicitly manage a context that applies to all elements of the
list. Let $\mathrm{List}(B)$ be the set of lists with elements from a
set $B$. If we have $A : \Set$ and $B : A \to \Set$ and
$X \in L(\Sigma a:A.\, B a)$, then we the resourced lists predicate
$\mathrm{RList}(X) \in L(\Sigma a : A.\, \mathrm{List}(B a))$ must
satisfy the equation:
\begin{displaymath}
  \mathrm{RList}(X) =
  \begin{array}[t]{l}
    \{ ((a, []), \alpha, (\synFalse, *)) \mid \alpha \in M \} \\
    \cup \\
    \{ \begin{array}[t]{@{}l}
         ((a, b :: bs), \alpha, (\synTrue, (v_1, v_2))) \mid \\
         \quad \exists \alpha_1, \alpha_2. 0 \leq M(\alpha, \alpha_1 \oplus \alpha_2) \land ((a, b), \alpha_1, v_1) \in X \land ((a, bs), \alpha_2, v_2) \in \mathrm{RList}(X) \}
       \end{array}
  \end{array}
\end{displaymath}
This equation has a least solution, by induction on the length of the
list being realised. This definition is somewhat involved, but in
essence states that a list is represented by tagged pairs, where
$\synFalse$ represents nil and $\synTrue$ represents cons, and that
the potential is distributed amongst the elements of the list as
needed.

Realisability of the nil and cons constructors, and of pattern
matching on lists is similar to case for booleans. We refer the reader
to the Agda development for the details.

\paragraph{Agda Formalisation} The construction of realisers for
booleans is carried out in the Agda module
\AgdaModule{AmortisedModel.Bool}. The constructions for lists with
case analysis are carried out in the module
\AgdaModule{AmortisedModel.List}.

\subsubsection{Realising QTT}

Summing up the previous sections, we have the following key theorem
paving the way to polytime soundness:
\begin{theorem}\label{thm:basic-qtt-realisability-model}
  For every resource monoid with a distinguished sub-monoid
  $(M, M_0)$, we obtain a realisability model of $\mathbb{N}$-QTT with
  $\Pi$ and $\Sigma$ types, a universe, extensional equality types,
  realisability reflection, and non-iterable datatypes.
\end{theorem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Realising Iteration for Implicit Polynomial Time}
\label{sec:realising-iteration}

The models constructed in the previous section only allow for
constant-time programs to be realised. To interpret the iterators of
the \emph{Cons-free} and LFPL-style systems, we need to use the
$\MaxPoly$ and $\PlusPoly$ resource monoids. We do this in this
section, where first we establish some operations that will be useful
to see how they capture the nesting of iterations inherent to polytime
computation.

\subsection{Iteration Resource Monoids}

To interpret iteration over a resource monoid $(M, M_0)$, we require
additional structure, which we call an \emph{Iteration Resource
  Monoid} to account for measurement of the sizes of iterable data
structures and the effects of iteration on potentials.

\subsubsection{Definition}
We require:
\begin{enumerate}
\item a function $\mathit{size} : \mathbb{N} \to M$ that gives the
  potential of an iterable data structure of a given
  size;
\item a function $\mathit{raise} : M \to M$ that raises the
  (polynomial) degree of some potential; and
\item a function $\mathit{scale} : \mathbb{N} \times M \to M$ that
  scales a potential for a fixed number of iterations.
\end{enumerate}
These functions must satisfy the following properties:
\begin{enumerate}
\item $M_0$ is closed under the $\mathit{raise}$ operation;
\item for all $\alpha$ and $n$,
  $0 \leq M(\mathit{raise}(\alpha) \oplus \mathit{size}(n),
  \mathit{scale}(n,\alpha) \oplus \mathit{size}(n))$; and
\item for all $\alpha \in M_0$ and $n$,
  $0 \leq M(\mathit{scale}(1 + n, \alpha), \alpha \oplus
  \mathit{scale}(n, \alpha))$.
\end{enumerate}
The first property states that $\mathit{raise}$ is suitable as
potential for whole programs2, meaning that it does not make any
requirements on the existence of iterable data. Note that we do
\emph{not} require $M_0$ to contain $\mathit{size}(n)$ -- programs
themselves may not contain iterable data, all potential for iteration
must be delivered externally. A useful intuition is that
$\mathit{scale}(n,\alpha)$ represents the potential required for at
most $n$ iterations that require potential $\alpha$, whereas
$\mathit{raise}(\alpha)$ represents the potential required for a
number of iterations that depends on the context. This is the
motivation behind the second required property, which states that
having $\mathit{raise}(\alpha)$ potential implies having
$\mathit{scale}(n,\alpha)$ potential when the current size is $n$. The
third property states that $\mathit{scale}$ decomposes as expected on
potentials that do not include any size potential.

Note that $\mathit{scale}(n,\alpha)$ is not the same as the action
$n \cdot \alpha$ defined in \autoref{sec:resource-monoids}. The latter
operation scales both size and function potential, but the former only
scales the function potential.

\subsubsection{Polynomial Iteration Resource Monoids}

Both of the polymonial resource monoids defined in
\defref{def:polynomial-resource-monoids} support the structure of an
Iteration Resource Monoid. We define:
\begin{enumerate}
\item $\mathit{size}(n) = (n, 0)$
\item $\mathit{raise}(n, p) = (n, xp)$
\item $\mathit{scale}(m, (n, p)) = (n, m \cdot p)$
\end{enumerate}
Note that $\mathit{raise}$ does indeed raise the degree of the
polynomial involved. Property 2 above is satisfied because for any
polynomial we have $(m \cdot p)(x) \leq (xp)(x)$ whenever $m \leq x$.

\subsubsection{Realising Iterable Natural Numbers}

For any natural number $n$, we define its representation as a value
$\mathrm{natValue}(n) \in \ValSet$ by recursion.
\begin{displaymath}
  \begin{array}{lcl}
    \mathrm{natValue}(0) &=& (\synTrue , *) \\
    \mathrm{natValue}(1 + n) &=& (\synFalse, \mathrm{natValue}(n))
  \end{array}
\end{displaymath}
This representation uses a tagged pair approach similar to our
representation of lists in \autoref{sec:noniter-model}. Using this, we
can define what it means for a natural number to be realisable via
$\mathrm{Nat} \in L(\mathbb{N})$:
\begin{displaymath}
  \mathrm{Nat} = \{(n, \alpha, \mathrm{natValue}(n)) \mid n \in \mathbb{N}, 0 \leq M(\alpha, \mathit{size}(n+1)) \}
\end{displaymath}
So a natural number $n$ is realised by the value
$\mathrm{natValue}(n)$ as long as we have at least
$\mathit{size}(n+1)$ potential (we add one to make the LFPL soundness
proof easier).  This gives us the ability to represent natural numbers
as a type in QTT, but in order to iterate (and construct in the case
of LFPL), we need to construct specific realisers for the
\emph{Cons-free} and LFPL systems.

\subsection{The Cons-Free System}
\label{sec:cons-free-sound}

The \emph{Cons-free} system uses the $\MaxPoly$ resource monoid, with
the distinguished sub-monoid being those elements that are $0$ in the
size component. We enumerate the features of the \emph{Cons-free}
system and justify their realisability with the $\MaxPoly$ resource
monoid:
\begin{enumerate}
\item Duplication of natural numbers by $\dupNat(M)$ is realisable by
  the expression $(0,0)$, which creates a pair by copying the input
  variable twice. By the cost semantics in
  \autoref{sec:machine-model}, this takes $1$ step of computation (we
  assume that it is actually implemented via some pointer copy).  The
  resource accounting for this realiser works because the size
  component required for the output is the maximum of the size
  components of the two elements, and since $n \sqcup n = n$, we have
  enough resources to fulfil this.
\item Construction of natural numbers is not realisable. In a putative
  $\conSucc$ rule, we would need to get an additional unit of size
  resource from nowhere.
\item Iteration is realised by constructing a realising expression in
  the expression language from the given expressions for the zero and
  successor cases that uses the in-built recursion of the
  language. The proof that resources are correctly accounted for is
  carried out by induction on the natural being iterated over. For
  $n$, we require potential
  $\mathit{scale}(n,\mathit{acct}(4) \rplus \gamma_{\mathit{succ}})
  \rplus (\mathit{acct}(2) \rplus \gamma_{\mathit{zero}})$, where
  $\gamma_{\mathit{succ}}$ and $\gamma_{\mathit{zero}}$ are the
  potentials required by the successor and zero cases respectively. By
  Property (2) of iteration resource monoids, above, we know that
  $\mathit{raise}(\mathit{acct}(4) \rplus \gamma_{\mathit{succ}})
  \rplus (\mathit{acct}(2) \rplus \gamma_{\mathit{zero}})$ always
  dominates this requirement when paired with the potential
  $\mathit{size}(n)$ from the input. Therefore, this latter
  expression, plus some administrative set up costs, is the required
  potential for the whole iterator.
\end{enumerate}
Together, we have a soundness result for the \emph{Cons-free} system,
that ensures that every term in the $\sigma = 1$ fragment is
realisable by a \emph{correct} program that terminates in polynomial
time for all inputs:
\begin{theorem}[Soundness for the Cons-free System]
  \label{thm:cons-free-soundness}
  If we have a term $n : \tyNat \vdash M \stackrel1: T(n)$ then there
  exists a realising expression $E$ and polynomial $p$ such that for
  all $n \in \mathbb{N}$, there exists $v \in \ValSet$ and
  $k \in \mathbb{N}$ such that
  $E, [\mathrm{natValue}(n)] \Downarrow_k v$, $k \leq p(n)$ and $v$ is
  a realising value for
  $\llbracket M \rrbracket(n) \in \llbracket X \rrbracket(n)$.
\end{theorem}

\paragraph{Agda Formalisation} The realisability of the
\emph{Cons-free} system iterator and the soundness property of the
whole system are formalised in the Agda modules \AgdaModule{ConsFree}
and \AgdaModule{ConsFree.Iterator}. The soundness theorem above is a
combination of the inequality definition in our indexed linear
preorder and the QTT model sketched in \autoref{sec:qtt-models}.

\subsection{The LFPL System}
\label{sec:lfpl-sound}

The LFPL system uses the $\PlusPoly$ resource monoid, with the
distinguished sub-monoid again being those elements that are $0$ in
the size component. With this resource monoid, the capabilities
offered at the QTT level are altered:
\begin{enumerate}
\item We can no longer duplicate natural numbers, because
  $\tyNat \otimes \tyNat$ requires twice as much size resource as
  $\tyNat$, due to the combining operation on size potentials being
  addition.
\item We define the realisability specification for diamonds
  $\Diamond \in L(1)$ to be:
  \begin{displaymath}
    \Diamond = \{ (*, \alpha, *) \mid 0 \leq M(\alpha, \mathit{size}(1)) \}
  \end{displaymath}
  Hence a diamond represents at least one unit of size resource,
  matching the intuitive explanation given in
  \autoref{sec:lfpl-intro}.
\item With this definition of realisability for $\Diamond$s, it is
  possible to realise the $\conZero$ and $\conSucc$ constructors for
  natural numbers. By the additive combination of size resources we
  get $1$ from the diamond and $n+1$ from the predecessor to total
  $n+2$ for a new number. Note that, even if we add a $\Diamond$ type
  to the \emph{Cons-free} system, it would still not be possible to
  realise the constructors, because we would only have
  $1 \sqcup (n+1) = n+1$ size resource for the output.
\item The construction of the recursor follows a very similar proof to
  the realisability of \emph{Cons-free} iterator, up to some
  additional work to make sure that the dummy $*$ values representing
  the diamond components end up in the right places. This additional
  work is revealed in the required potential for the LFPL iterator
  being
  $\mathit{raise}(\mathit{acct}(8) \rplus \gamma_{\mathit{succ}})
  \rplus (\mathit{acct}(2) \rplus \gamma_{\mathit{zero}})$, so
  slightly higher in the successor case.
\end{enumerate}

Soundness for the LFPL system is similar to the \emph{Cons-free}
system, except for a $+1$ to the input to the polynomial, to account
for the fact that we cost one size unit for the $\conZero$
constructor.
\begin{theorem}[Soundness for the LFPL-style System]
  \label{thm:lfpl-soundness}
  If we have a term $n : \tyNat \vdash M \stackrel1: T(n)$ then there
  exists a realising expression $E$ and polynomial $p$ such that for
  all $n \in \mathbb{N}$, there exists $v \in \ValSet$ and
  $k \in \mathbb{N}$ such that
  $E, [\mathrm{natValue}(n)] \Downarrow_k v$, $k \leq p(n+1)$ and $v$
  is a realising value for
  $\llbracket M \rrbracket(n) \in \llbracket X \rrbracket(n)$.
\end{theorem}

The proofs of well-accounted realisability for the LFPL iterator, and
the \emph{Cons-free} iterator, could be adapted to any other
inductively defined type that is finitely branching. This is not
immediately necessary, as evidenced by the construction of other
datatypes in \autoref{sec:prog-datatypes}. Nevertheless, native tree
type where the iterability is proportional to the total number of
nodes would be useful.

\paragraph{Agda Formalisation} The realisability of the LFPL system
iterator and the soundness property of the whole system are formalised
in the Agda modules \AgdaModule{LFPL} and \AgdaModule{LFPL.Iterator}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related and Future Work}
\label{sec:conclusion}

We have presented an two extensions of Quantitative Type Theory that
soundly and completely captures polynomial time.

\subsection{Systems for Polytime}

BLL, SAL, LAL

THe ones from Fossacs 2022, and POPL23(?) Romain Per...

Neil Jones

\subsection{Other Complexity Classes}

\begin{enumerate}
\item Lower complexity classes, like LOGSPACE
\item Higher complexity classes, like EXPTIME, ELEMENTARY
\end{enumerate}

\subsection{Explicit Resource Accounting}

\begin{enumerate}
\item RAML, Hofmann-Jost, etc.
\item lambda-amor
\item
\end{enumerate}

\subsection{Linear and Substructural Dependent Types}

Linear LF

Pradic et al

Grtt

Penn

Abel

\subsection{Implementation} Could use Idris, but it has the wrong
semiring, and its datatype definition facility is too permissive.

\subsection{Towards Synthetic Computational Complexity Theory?}

Blah

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
  FIXME
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

\end{document}
\endinput